decompr: GVC decomposition in R

### 2015-02-09 draft.Rnw
Update draft.Rnw
Final changes and additions

\vspace{\stretch{1}}
\abstract{Global Value Chains have become a central unit of analysis in research on international trade.
However, the complex matrix transformations at the basis of most Value Chain indicators still constitute a significant entry barrier to
the field.
The R package \textbf{decompr} solves this problem by implementing the algorithms for the analysis of Global Value Chains as R procedures, thereby simplifying the decomposition process..
Two methods for gross export flow decomposition using Inter-Country Input-Output tables are provided.
The first method is the \textbf{Wang-Wei-Zhu} algorithm which splits bilateral gross exports into 16 value added components. These components can broadly be divided into domestic and foreign value added in exports.
The second method concerns a decomposition based on the classical \textcite{wale36} insight.
It derives the value added origins of an industry's exports by source country and source industry, using easily available
gross trade data.
This article summarises the methodology of the algorithms, describes the format of the input and output data, 
and exemplifies the usefulness of the two methods on the basis of a simple example data set.}
\pagebreak

\section{Introduction}
\epigraph{Hello?\\...\\ is it me you're looking for?}{Lionel Richie}
Global Value Chains (GVCs) refer to the quickly expanding internationalization of production networks.
Most goods we use nowadays consist of parts that are sourced from different corners of the planet 
and are assembled across different continents. A popular example of this development is the Iphone,
which uses inputs from at least five countries (USA, China, Germany, Taiwan, South Korea) and is assembled in two (USA and China).
This has made GVCs a central topic in research on trade and development policy.
Both policy makers and academia increasingly value the growth opportunities GVCs offer to global trade and, 
especially, to developing countries.
However, analysing this phenomenon empirically requires complex matrix manipulations, 
since the relevant data is only available in the form of gross flows.
This package enables researchers with little background in matrix algebra and linear programming to easily derive standard GVC indicators for statistical analysis.

The \textbf{decompr} package uses Inter-Country Input-Output tables (ICIOs), such as those published by the OECD and WTO (TiVA), % CITE!! 
the World Input Output Database \parencite{timmer2012world},  or national statistics bureaus, as input.
These tables state supply and demand relationships in gross terms between industries within and across countries.
For instance, let us look at the example of the leather used in German manufactured car seats.
The ICIOs quantify the value of inputs that the Turkish leather and textiles industry
supplies to the German transport equipment industry.

The problem of these tables measuring gross trade flows, 
is that they do not reveal how much of the value was added in the supplying industry,
and how much of the value was added in previous stages of production, performed by other industries or even countries.
The \textbf{Leontief} decomposition of gross trade flows solves this problem by reallocating the value of intermediate goods used by industries to the original producers.
In our example, the use of Argentinian agricultural produce (raw hides) is subtracted from the Turkish leather industry and added to the Argentinian agricultural industry.

The \textbf{Wang-Wei-Zhu} (henceforth WWZ) decomposition goes a step further by not only revealing the source of the value added,
but also breaking down exports into different categories, according to final usage and destination.
It implements the theoretical work of \textcite{wang2014quantifying}.
The main categories in this framework are listed below.

@@

Trade flow analysis often involves studying development of a certain variable (set) over time, thus taking the panel form.
However, no existing trade flow analysis explicitly exploints the time dimsion, making the analysis effectively a repeated cross-section.
As a design decision, the time dimension is therefore not implemented in the package itself.
In stead, we provide examples of how the repeated analysis can be easily implemented using decompr.

@@

\section{Data}
\label{sec:data}
Two data sets are included in the package, one real world data set, and one minimal data set (fictional), for training surposes.
The former is the WIOD regional Inter-Country Input Output tables for the year 2011 \parencite{timmer2012world}. 
The later, is a fictional 3-country 3-sector data set, which we will use throughout this article to demonstrate the usage of the decompr package.

@@

The data is set up in order to illustrate the benefits of the decompositions.
We do this by following the flows of intermediate goods through a fictional GVC and showing how the readily available gross trade flows differ from the decomposed value added flows.
To this end, we construct the elements of the input-output tables such that we have two countries and two industries that focus on upstream tasks, which means they focus on supplying other industries, and one industry and one country that is specialized in downstream tasks,i.e.~it serves mainly final demand.
In our example the upstream industries are Agriculture and Textiles while the downstream industry in Transport Equipment.
Similarly, Argentina and Turkey represent upstream countries with Germany being located downstream within this specific value chain (see \cref{tab:leather}.

@@

The first step of the analytic process is to load the input object and create a \textbf{decompr} class object which contains the data structures for the decompositions.
This step is not need when using the \textbf{decomp} wrapper function, more on this later.

@@

As can be seen above, a \textbf{decompr} class object is in fact a list containing thirty different objects.
For example, \textbf{Eint} is an object that collects the intermediate goods exports of the industries,
while \textbf{G} refers to the number of countries that are analysed.
Depending on the choice of decomposition, all or some of these objects are used.

\section{Leontief decomposition}
\label{sec:leontief}
Let us now turn to the algorithms, starting with the \textbf{Leontief} decomposition. We shortly describe the theoretical
derivation of the method to expose the internal steps of the \textbf{decompr} package. After which, we turn to the technical implementation.
Finally, we describe the output.

@@

\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textbf{V} is a N x N matrix with the diagonal representing the direct value added contribution of the industries,
\textbf{A} is the Input-Output coefficient matrix with dimension N x N, i.e. it gives the direct input flows between
industries required for 1\$ of output, $B = (I-A)^{-1}$ is then the so called Leontief inverse, and \textbf{T} indicates
a matrix transpose operation.
\textbf{VB} gives thus a N x N matrix of so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries.
Looking from the perspective of the supplying industries, the matrix gives the value added that they contribute to the using industry's production.
If we multiply it with a N x N matrix whose diagonal specifies each industry's total output, we get value added origins as absolute values instead of shares.

The application of the Leontief insight to ICIOs as opposed to national Input-Output tables for our source decomposition
is straightforward.
\textbf{V} refers now to a vector of direct value added contribution of all industries across the different countries. Its dimension is correspondingly 1 x GN.
\textbf{A} is now of dimension GN x GN and gives the industry flows including cross border relationships.
Since we are interested in the value added origins of exports we multiply these two matrices with a GN x GN matrix whose diagonal we fill with each industry's exports, $E$, such that the basic equation behind the source decomposition is given by $V(I-A)^{-1}E$.
In a simple example with two countries (\textbf{k} and \textbf{l}) and industries (\textbf{i} and \textbf{j}) we can zoom in to see the matrices' content:

\begin{align}
\begin{split}
V (I- A &)^{-1} E=
\begin{pmatrix}
v_{k}^{i}& 0& 0& 0\\
0& v_{k}^{j}& 0& 0\\
0& 0& v_{l}^{i}& 0\\
0& 0& 0& v_{l}^{j}
\end{pmatrix}
*
\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
*
\begin{pmatrix}
e_{k}^{i}& 0& 0& 0\\
0& e_{k}^{j}& 0& 0\\
0& 0& e_{l}^{i}& 0\\
0& 0& 0& e_{l}^{j}
\end{pmatrix}
=
\\
&
\begin{pmatrix}
v_{k}^{i}b_{kk}^{ii}e_{k}^{i}& v_{k}^{i} b_{kk}^{ij}e_{k}^{j}& v_{k}^{i}b_{kl}^{ii}e_{l}^{i}& v_{k}^{i}b_{kl}^{ij}e_{l}^{j}\\
v_{k}^{j}b_{kk}^{ji}e_{k}^{i}& v_{k}^{j}b_{kk}^{jj}e_{k}^{j}& v_{k}^{j}b_{kl}^{ji}e_{l}^{i}& v_{k}^{j}b_{kl}^{jj}e_{l}^{j}\\
v_{l}^{i}b_{lk}^{ii}e_{k}^{i}& v_{l}^{i}b_{lk}^{ij}e_{k}^{j}& v_{l}^{i}b_{ll}^{ii}e_{l}^{i}& v_{l}^{i}b_{ll}^{ij}e_{l}^{j}\\
v_{l}^{j}b_{lk}^{ji}e_{k}^{i}& v_{l}^{j}b_{lk}^{jj}e_{k}^{j}& v_{l}^{j}b_{ll}^{ji}e_{l}^{i}& v_{l}^{j}b_{ll}^{jj}e_{l}^{j}
\end{pmatrix}
=
\begin{pmatrix}
vae_{kk}^{ii}& vae_{kk}^{ij}& vae_{kl}^{ii}& vae_{kl}^{ij}\\
vae_{kk}^{ji}& vae_{kk}^{jj}& vae_{kl}^{ji}& vae_{kl}^{jj}\\
vae_{lk}^{ii}& vae_{lk}^{ij}& vae_{ll}^{ii}& vae_{ll}^{ij}\\
vae_{lk}^{ji}& vae_{lk}^{jj}& vae_{ll}^{ji}& vae_{ll}^{jj}
\end{pmatrix}
\end{split}
\end{align}
\begin{align*}
&v_{c}^{s} = \frac{va_{c}^{s}}{y_{c}^{s}} = 1 - a_{kc}^{is} - a_{kc}^{js} - a_{lc}^{js} - a_{lc}^{is} \hspace{2cm} (c \in k,l \hspace{.5cm} s \in i,j),\\
&\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
=
\begin{pmatrix}
1-a_{kk}^{ii}& -a_{kk}^{ij}& -a_{kl}^{ii}& -a_{kl}^{ij}\\
-a_{kk}^{ji}& 1-a_{kk}^{jj}& -a_{kl}^{ji}& -a_{kl}^{jj}\\
-a_{lk}^{ii}& -a_{lk}^{ij}& 1-a_{ll}^{ii}& -a_{ll}^{ij}\\
-a_{lk}^{ji}& -a_{lk}^{jj}& -a_{ll}^{ji}& 1-a_{ll}^{jj}
\end{pmatrix}^{-1},
\end{align*}
and
\begin{equation*}
a_{cf}^{su} = \frac{inp_{cf}^{su}}{y_{f}^{u}}  \hspace{2cm} (c,f \in k,l \hspace{.5cm} s,u \in i,j),
\end{equation*}
where \(v_{s}^{c}\) gives the share of industry \emph{s}'s value added, \(va_{c}^{s},\) in output, \(y_{s}^{c}\), and \(e_{k}^{i}\) indicates gross exports. Finally, \(a_{su}^{cf}\) denotes the share of inputs, \(inp_{su}^{cf}\), in output.
The elements of the $V(I-A)^{-1}E$ or $vae$ matrix are our estimates for the country-industry level value added
origins of each country-industry's exports.
\textbf{decompr} implements this algorithm into R to automate the process of deriving the matrix. 
Equipped with it, researchers can calculate standard GVC indicators.
Examples include \citet{dahuetal01}'s Vertical Specialisation ratio at the industry-level, 
which can be derived by summing for each country and industry across the value added of all foreign countries and industries,
and \citet{rojoguno12}'s so-called VAX ratio.
Alternatively, the four dimensions of the matrix (source country, source industry, using country, using industry) allow for industry-level gravity-type estimations of value added trade flows.

@@

\section{Wang-Wei-Zhu decomposition}
\label{sec:wwz}
The Wang-Wei-Zhu decomposition builds upon the Leontief insight but uses, in addition, further valuable information
provided in ICIOS.
More specifically, the Leontief decomposition traces the value added back to where it originates but ICIOS also contain data on how the value added is subsequently used.
This information is extracted by the Wang-Wei-Zhu decomposition, which thereby allows a much more detailed look at the structures of international production networks and the respective positions of countries and industries within them.

@@

This is why we present here only the final equation for a two country one industry model (equation 22 in WWZ) and refer the interested reader to the original paper by Wang, Wei, and Zhu (2014).

@@

where $F^{kl}$ is the final demand in $l$ for goods of $k$ and $L^{ll}$ refers to the national Leontief inverse as opposed to the Inter-Country inverse $B$. As can be seen from equation (\ref{eq:wwz}), the Wang-Wei-Zhu decomposition splits gross exports into 16 terms with three main categories given by domestic value added in exports (\textbf{DViX\_B}), foreign value added in exports (\textbf{FVA}), and purely double counted terms (\textbf{PDC}). The main categories are further divided according to their final destination so that the final decomposition is given by:

@@

\item Domestic Value Added absorbed abroad (\textbf{VAX\_G}, T1-5)

@@

\item Domestic Value added in final exports (\textbf{DVA\_FIN}, T1)
\item Domestic Value added in intermediate exports

@@

\item Domestic Value added in intermediate exports absorbed by direct importers (\textbf{DVA\_INT}, T2)
\item Domestic Value added in intermediate exports re-exported to third countries (\textbf{DVA\_INTrex}, T3-5)

@@

\item Domestic Value added in intermediate exports re-exported to third countries as intermediate goods to produce domestic final goods (\textbf{DVA\_INTrexI1}, T3)
\item Domestic Value added in intermediate exports re-exported to third countries as  final goods (\textbf{DVA\_INTrexF}, T4)
\item Domestic Value added in intermediate exports re-exported to third countries as intermediate goods to produce exports (\textbf{DVA\_INTrexI2}, T5)

@@

\item Domestic Value Added returning home (\textbf{RDV\_B}, T6-8)

@@

\item Domestic Value Added returning home as final goods (\textbf{RDV\_FIN}, T6)
\item Domestic Value Added returning home as final goods through third countries (\textbf{RDV\_FIN2}, T7)
\item Domestic Value Added returning home as intermediate goods (\textbf{RDV\_INT}, T8)

@@

\item Foreign Value added (\textbf{FVA}, T11-12/14-15 )

@@

\item Foreign Value added in final good exports (\textbf{FVA\_FIN}, T11/14)

@@

\item Foreign Value added in final good exports sourced from direct importer (\textbf{MVA\_FIN}, T11)
\item Foreign Value added in final good exports sourced from other countries (\textbf{OVA\_FIN}, T14)

@@

\item Foreign Value added in intermediate good exports (\textbf{FVA\_INT}, T12/15)

@@

\item Foreign Value added in intermediate good exports sourced from direct importer (\textbf{MVA\_INT}, T12)
\item Foreign Value added in intermediate good exports sourced from other countries(\textbf{OVA\_INT}, T15)

@@

\item Pure Double Counting (\textbf{PDC}, T9-10/13/16)

@@

The higher resolution of the WWZ decomposition comes at the cost of a lower dimension (source country, using country, using industry) since the current, highly aggregated, ICIOs render a four-dimensional decomposition unfeasible. This means that the two methods are complementary and imply a trade-off between detail and disaggregation.

@@

\subsection{Output}
The output when using the WWZ algorithm is a matrix with dimensions GNGx19 whereby 19 is the 16 objects the WWZ algorithm decomposes exports into, plus three checksums. GNG represents source country, source industry and using country whereas these terms are slightly ambiguous here due to the complex nature of the decomposition. More specifically, the using country can also be the origin of the foreign value added in the exports of the source country to the using country (see for example T11 and T12). Therefore we use the terms exporter, exporting industry, and direct importer instead. This becomes clearer when we take a look at specific examples.

@@

The first eight components relate to domestic value added of the exporting country contained in the sectoral exports of the exporting industry to the direct importer. For instance, the first non-zero element in \cref{tab:wwz} refers to $DVA\_FIN$ or domestic value added in final good exports. It shows that there are 5.47 units of domestic value added in the exports of final goods from Argentina's Agriculture industry to Turkey. In the same row the third term,  $DVA\_INTrexI1$, is slightly more complicated. As mentioned above it gives the amount of domestic Value added in intermediate exports re-exported to third countries as intermediate goods to produce domestic final goods. In our example this means that there are 1.14 units of domestic value added in the intermediate exports of Argentina's Agriculture industry to Turkey that are re-exported by Turkey as intermediates to a third country which produces final goods with it. Terms six to eight concern domestic value added that eventually returns home. $RDV\_FIN2$ reveals for example that there are 0.35 units of domestic value added in the intermediate exports of Argentina's Agriculture industry to Turkey that Turkey re-exports as intermediates to Argentina for the latter's final good production.

@@

Terms 13 to 16 collect the double counting of gross trade statistics that occurs when goods cross borders multiple times. $DDC$ captures double counting due to domestic value added, which is further classified according to the type of the ultimate export (final vs intermediate good). $MDC$ and $ODC$ on the other hand capture double counting due to foreign value added from either the direct importer or other countries. For the Argentina-Turkey case this implies for instance that there are 0.18 units of value added in the intermediate exports of Turkey to Argentina which are re-exported by Argentina's Agriculture industry to Turkey as intermediates and then again re-exported. Since they would be part of $MVA$ twice, they are now counted once as double-counted term.

>>>>

\vspace{\stretch{1}}
\abstract{Global Value Chains have become a central unit of analysis in research on international trade.
However, the complex matrix transformations at the basis of most Value Chain indicators still constitute a significant entry barrier to
the field.
The R package \textbf{decompr} solves this problem by implementing the algorithms for the analysis of Global Value Chains as R procedures,
thereby simplifying the decomposition process.
Two methods for gross export flow decomposition using Inter-Country Input-Output tables are provided.
The first method concerns a decomposition based on the classical \textbf{\textcite{wale36}} insight.
It derives the value added origins of an industry's exports by source country and source industry, using easily available
gross trade data.
The second method is the \textbf{Wang-Wei-Zhu} algorithm which splits bilateral gross exports into 16 value added components. 
These components can broadly be divided into domestic and foreign value added in exports.
Using the results of the two decompositions, \textbf{decompr} provides a set of standard Global Value Chain indicators, 
such as the now standard Vertical Specialisation ratio.
This article summarises the methodology of the algorithms, describes the format of the input and output data, 
and exemplifies the usefulness of the two methods on the basis of a simple example data set.}
\pagebreak

\section{Introduction}
\epigraph{Hello?\\...\\ is it me you're looking for?}{Lionel Richie}
Global Value Chains (GVCs) refer to the quickly expanding internationalization of production networks.
Most goods we use nowadays consist of parts that are sourced from different corners of the planet 
and are assembled across different continents. A popular example of this development is the Iphone,
which uses inputs from at least five countries (USA, China, Germany, Taiwan, South Korea) and is assembled in two (USA and China).
This has made GVCs a central topic in research on trade and development policy.
Both policy makers and academia increasingly value the growth opportunities GVCs offer to global trade and, 
especially, to developing countries.
However, analysing this phenomenon empirically requires complex matrix manipulations, 
since the relevant data is only available in the form of gross flows.
The \textbf{decompr} package enables researchers with little background in matrix algebra and linear programming to easily derive standard GVC indicators for statistical analysis.

The package uses Inter-Country Input-Output tables (ICIOs), such as those published by the OECD and WTO (TiVA), % CITE!! 
the World Input Output Database \parencite{timmer2012world},  or national statistics bureaus, as input.
These tables state supply and demand relationships in gross terms between industries within and across countries.
For instance, let us look at the example of the leather used in German manufactured car seats.
The ICIOs quantify the value of inputs that the Turkish leather and textiles industry
supplies to the German transport equipment industry. The problem of these tables measuring gross trade flows, 
is that they do not reveal how much of the value was added in the supplying industry,
and how much of the value was added in previous stages of production, performed by other industries or even countries.

The \textbf{Leontief} decomposition of gross trade flows solves this problem by reallocating the value of intermediate goods used by industries to the original producers.
In our example, the use of Argentinian agricultural produce (raw hides) is subtracted from the Turkish leather industry and added to the Argentinian agricultural industry.
The \textbf{Wang-Wei-Zhu} (henceforth WWZ) decomposition goes a step further by not only revealing the source of the value added,
but also breaking down exports into different categories, according to final usage and destination.
It implements the theoretical work of \textcite{wang2014quantifying}.
The main categories in this framework are listed below.

@@

Trade flow analysis often involves studying the development of a certain variable (set) over time, thus taking the panel form.
However, no existing trade flow analysis explicitly exploits the time dimension, making the analysis effectively a repeated cross-section.
As a design decision, the time dimension is therefore not implemented in the package itself.
Instead, we provide examples of how the repeated analysis can be easily implemented using decompr.

@@

\section{Data}
\label{sec:data}
Two data sets are included in the package, one real world data set and one minimal data set (fictional) for demonstration purposes.
The former is the WIOD regional Inter-Country Input Output tables for the year 2011 \parencite{timmer2012world}. 
The latter is a fictional 3-country 3-sector data set, which we will use throughout this article to demonstrate the usage and advantages of the decompr package.

@@

This data is set up in order to illustrate the benefits of the decompositions.
We do this by following the flows of intermediate goods through a fictional GVC and by showing how the readily available gross trade flows differ from the decomposed value added flows.
To this end, we construct the elements of the input-output tables such that we have two countries and two industries that focus on upstream tasks, 
which means they focus on supplying other industries,
and one country and industry that is specialized in downstream tasks, i.e.~it serves mainly final demand.
In our example the upstream industries are Agriculture and Textiles while the downstream industry is Transport Equipment.
Similarly, Argentina and Turkey represent upstream countries with Germany being located downstream within this specific value chain (see \cref{tab:leather}).

@@

The first step of the analytic process is to load the input object and create a \textbf{decompr} class object, 
which contains the data structures for the decompositions.
This step is not needed when using the \textbf{decomp} wrapper function but more on this later.

@@

As can be seen above, a \textbf{decompr} class object is in fact a list containing thirty different objects.
For example, \textbf{Eint} is an object that collects the intermediate goods exports of the industries,
while \textbf{G} refers to the number of countries that are analysed.
Depending on the choice of the decomposition, all or some of these objects are used.

\section{Leontief decomposition}
\label{sec:leontief}
Let us now turn to the algorithms, starting with the \textbf{Leontief} decomposition. 
We shortly describe the theoretical derivation of the method to expose the internal steps of the \textbf{decompr} package. 
Afterwards, we turn to the technical implementation and, finally, we describe the output.

@@

\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textbf{V} is a N x N matrix with the diagonal representing the direct value added contribution of N industries,
\textbf{A} is the Input-Output coefficient matrix with dimension N x N, i.e. it gives the direct input flows between
industries required for 1\$ of output, and $B = (I-A)^{-1}$ is the so called Leontief inverse.
\textbf{VB} gives thus a N x N matrix of so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries.
Looking from the perspective of the supplying industries, the matrix gives the value added that they contribute to the using industry's production.
If we multiply it with a N x N matrix whose diagonal specifies each industry's total output or exports, we get value added origins as absolute values instead of shares.

The application of the Leontief insight to ICIOs as opposed to national Input-Output tables for our \textbf{Leontief} decomposition is straightforward.
\textbf{V} refers now to a vector of direct value added contribution of all industries across the different countries. 
Its dimension is correspondingly 1 x GN, where G is according to the number of countries.
\textbf{A} is now of dimension GN x GN and gives the industry flows including cross border relationships.
Since we are interested in the value added origins of exports we multiply these two matrices with a GN x GN matrix whose diagonal we fill with each industry's exports, $E$, 
such that the basic equation behind the source decomposition is given by $V(I-A)^{-1}E$ 
\footnote{When using the textbf{leontief_output} function, the value added multiplier is instead multiplied with each industry's output.}.
In a simple example with two countries (\textbf{k} and \textbf{l}) and industries (\textbf{i} and \textbf{j}) we can zoom in to see the matrices' content:

\begin{align}
\begin{split}
V (I- A &)^{-1} E=
\begin{pmatrix}
v_{k}^{i}& 0& 0& 0\\
0& v_{k}^{j}& 0& 0\\
0& 0& v_{l}^{i}& 0\\
0& 0& 0& v_{l}^{j}
\end{pmatrix}
*
\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
*
\begin{pmatrix}
e_{k}^{i}& 0& 0& 0\\
0& e_{k}^{j}& 0& 0\\
0& 0& e_{l}^{i}& 0\\
0& 0& 0& e_{l}^{j}
\end{pmatrix}
=
\\
&
\begin{pmatrix}
v_{k}^{i}b_{kk}^{ii}e_{k}^{i}& v_{k}^{i} b_{kk}^{ij}e_{k}^{j}& v_{k}^{i}b_{kl}^{ii}e_{l}^{i}& v_{k}^{i}b_{kl}^{ij}e_{l}^{j}\\
v_{k}^{j}b_{kk}^{ji}e_{k}^{i}& v_{k}^{j}b_{kk}^{jj}e_{k}^{j}& v_{k}^{j}b_{kl}^{ji}e_{l}^{i}& v_{k}^{j}b_{kl}^{jj}e_{l}^{j}\\
v_{l}^{i}b_{lk}^{ii}e_{k}^{i}& v_{l}^{i}b_{lk}^{ij}e_{k}^{j}& v_{l}^{i}b_{ll}^{ii}e_{l}^{i}& v_{l}^{i}b_{ll}^{ij}e_{l}^{j}\\
v_{l}^{j}b_{lk}^{ji}e_{k}^{i}& v_{l}^{j}b_{lk}^{jj}e_{k}^{j}& v_{l}^{j}b_{ll}^{ji}e_{l}^{i}& v_{l}^{j}b_{ll}^{jj}e_{l}^{j}
\end{pmatrix}
=
\begin{pmatrix}
vae_{kk}^{ii}& vae_{kk}^{ij}& vae_{kl}^{ii}& vae_{kl}^{ij}\\
vae_{kk}^{ji}& vae_{kk}^{jj}& vae_{kl}^{ji}& vae_{kl}^{jj}\\
vae_{lk}^{ii}& vae_{lk}^{ij}& vae_{ll}^{ii}& vae_{ll}^{ij}\\
vae_{lk}^{ji}& vae_{lk}^{jj}& vae_{ll}^{ji}& vae_{ll}^{jj}
\end{pmatrix}
\end{split}
\end{align}
\begin{align*}
&v_{c}^{s} = \frac{va_{c}^{s}}{y_{c}^{s}} = 1 - a_{kc}^{is} - a_{kc}^{js} - a_{lc}^{js} - a_{lc}^{is} \hspace{2cm} (c \in k,l \hspace{.5cm} s \in i,j),\\
&\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
=
\begin{pmatrix}
1-a_{kk}^{ii}& -a_{kk}^{ij}& -a_{kl}^{ii}& -a_{kl}^{ij}\\
-a_{kk}^{ji}& 1-a_{kk}^{jj}& -a_{kl}^{ji}& -a_{kl}^{jj}\\
-a_{lk}^{ii}& -a_{lk}^{ij}& 1-a_{ll}^{ii}& -a_{ll}^{ij}\\
-a_{lk}^{ji}& -a_{lk}^{jj}& -a_{ll}^{ji}& 1-a_{ll}^{jj}
\end{pmatrix}^{-1},
\end{align*}
and
\begin{equation*}
a_{cf}^{su} = \frac{inp_{cf}^{su}}{y_{f}^{u}}  \hspace{2cm} (c,f \in k,l \hspace{.5cm} s,u \in i,j),
\end{equation*}
where \(v_{s}^{c}\) gives the share of industry \emph{s}'s value added, \(va_{c}^{s},\) in output, \(y_{s}^{c}\), and \(e_{k}^{i}\) indicates gross exports. Finally, \(a_{su}^{cf}\) denotes the share of inputs, \(inp_{su}^{cf}\), in output.
The elements of the $V(I-A)^{-1}E$ or $vae$ matrix are our estimates for the country-industry level value added
origins of each country-industry's exports.
\textbf{decompr} implements this algorithm into R to automate the process of deriving the matrix. 
Equipped with it, researchers can calculate standard GVC indicators.
Examples include \citet{dahuetal01}'s Vertical Specialisation ratio at the industry-level using the textbf{vertical\_specialisation} function, 
which sums for each country and industry across the value added of all foreign countries and industries,
and \citet{rojoguno12}'s so-called VAX ratio.
Alternatively, the four dimensions of the matrix (source country, source industry, using country, using industry) allow for industry-level gravity-type estimations of value added trade flows.

@@

\section{Wang-Wei-Zhu decomposition}
\label{sec:wwz}
The Wang-Wei-Zhu decomposition builds upon the Leontief insight but uses, in addition, further valuable information provided in ICIOs.
More specifically, the Leontief decomposition traces the value added back to where it originates but ICIOs also contain data on how the value added is subsequently used.
This information is extracted by the Wang-Wei-Zhu decomposition, 
which thereby allows a much more detailed look at the structures of international production networks and the respective positions of countries and industries within them.

@@

This is why we present here only the final equation for a two country one industry model (equation 22 in WWZ) and refer the interested reader to the original paper by \textcite{wang2014quantifying}.

@@

where $F^{kl}$ is the final demand in $l$ for goods of $k$, 
$L^{ll}$ refers to the national Leontief inverse as opposed to the Inter-Country inverse $B$, 
and \textbf{T} indicates a matrix transpose operation. 
As can be seen from equation (\ref{eq:wwz}), 
the Wang-Wei-Zhu decomposition splits gross exports into 16 terms with three main categories given by domestic value added in exports (\textbf{DViX\_B}), foreign value added in exports (\textbf{FVA}), and purely double counted terms (\textbf{PDC}). 
The main categories are further divided according to their final destination so that the final decomposition is given by:

@@

\item Domestic value added absorbed abroad (\textbf{VAX\_G}, T1-5)

@@

\item Domestic value added in final exports (\textbf{DVA\_FIN}, T1)
\item Domestic value added in intermediate exports

@@

\item Domestic value added in intermediate exports absorbed by direct importers (\textbf{DVA\_INT}, T2)
\item Domestic value added in intermediate exports re-exported to third countries (\textbf{DVA\_INTrex}, T3-5)

@@

\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce domestic final goods (\textbf{DVA\_INTrexI1}, T3)
\item Domestic value added in intermediate exports re-exported to third countries as  final goods (\textbf{DVA\_INTrexF}, T4)
\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce exports (\textbf{DVA\_INTrexI2}, T5)

@@

\item Domestic value added returning home (\textbf{RDV\_B}, T6-8)

@@

\item Domestic value added returning home as final goods (\textbf{RDV\_FIN}, T6)
\item Domestic value added returning home as final goods through third countries (\textbf{RDV\_FIN2}, T7)
\item Domestic value added returning home as intermediate goods (\textbf{RDV\_INT}, T8)

@@

\item Foreign value added (\textbf{FVA}, T11-12/14-15 )

@@

\item Foreign value added in final good exports (\textbf{FVA\_FIN}, T11/14)

@@

\item Foreign value added in final good exports sourced from direct importer (\textbf{MVA\_FIN}, T11)
\item Foreign value added in final good exports sourced from other countries (\textbf{OVA\_FIN}, T14)

@@

\item Foreign value added in intermediate good exports (\textbf{FVA\_INT}, T12/15)

@@

\item Foreign value added in intermediate good exports sourced from direct importer (\textbf{MVA\_INT}, T12)
\item Foreign value added in intermediate good exports sourced from other countries(\textbf{OVA\_INT}, T15)

@@

\item Pure double counting (\textbf{PDC}, T9-10/13/16)

@@

The higher resolution of the WWZ decomposition comes at the cost of a lower dimension (source country, using country, using industry) since the current, highly aggregated, ICIOs render a four-dimensional decomposition unfeasible.
This means that the two methods are complementary and imply a trade-off between detail and disaggregation.

@@

\subsection{Output}
The output when using the WWZ algorithm is a matrix with dimensions GNGx19, 
whereby 19 consists of the 16 objects the WWZ algorithm decomposes exports into, plus three checksums. 
GNG represents source country, source industry and using country whereas these terms are slightly ambiguous here due to the complex nature of the decomposition. 
More specifically, the using country can also be the origin of the foreign value added in the exports of the source country to the using country (see for example T11 and T12). 
Therefore we use the terms exporter, exporting industry, and direct importer instead. 
This becomes much clearer when we take a look at specific examples.

@@

The first eight components relate to domestic value added of the exporting country contained in the sectoral exports of the exporting industry to the direct importer. 
For instance, the first non-zero element in \cref{tab:wwz} refers to $DVA\_FIN$, or domestic value added in final good exports. 
It shows that there are 5.47 units of domestic value added in the exports of final goods from Argentina's Agriculture industry to Turkey. 
In the same row the third term,  $DVA\_INTrexI1$, is slightly more complicated. 
As mentioned above it gives the amount of domestic Value added in intermediate exports re-exported to third countries as intermediate goods to produce domestic final goods. 
In our example this means that there are 1.14 units of domestic value added in the intermediate exports of Argentina's Agriculture industry to Turkey, 
that are re-exported by Turkey as intermediates to a third country which produces final goods with it. 
Terms six to eight concern domestic value added that eventually returns home. 
$RDV\_FIN2$ reveals for example that there are 0.35 units of domestic value added in the intermediate exports of Argentina's Agriculture industry to Turkey, 
that Turkey re-exports as intermediates to Argentina for the latter's final good production.

@@

Terms 13 to 16 collect the double counting of gross trade statistics that occurs when goods cross borders multiple times. 
$DDC$ captures double counting due to domestic value added, 
which is further classified according to the type of the ultimate export (final vs intermediate good). 
$MDC$ and $ODC$ on the other hand capture double counting due to foreign value added from either the direct importer or other countries. 
For the Argentina-Turkey case this implies, for instance, that there are 0.18 units of value added in the intermediate exports of Turkey to Argentina which are re-exported by Argentina's Agriculture industry to Turkey as intermediates and then again re-exported. 
Since they would be part of $MVA$ twice, they are now counted once as double-counted term.
